---
title: "DOC Trend Exploration"
author: "Nick Gubbins"
output: html_document
editor_options: 
  chunk_output_type: console
toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      include = TRUE,
                      tidy.opts = list(width.cutoff = 60), 
                      #tidy = TRUE,
                      warning = F,
                      message = F)
# needed packages
library(here)
library(tidyverse)
library(lubridate)
library(mapview)
mapviewOptions(fgb = FALSE, georaster = FALSE)
library(sf)
library(tigris)
library(RColorBrewer)
library(leaflet)
library(sp)
library(ggthemes)
library(macrosheds)
library(tidyr)
library(trend)
library(feather)
library(lfstat)
library(mapview)

# set color ramps
colors = colorRampPalette(c('blue', 'dark red'))
```
# Concentration

## Set up MS data

First find all data downloaded to MS directory.

```{r}
macrosheds_root <- here('ms_data')


site_data <- ms_load_sites()
domain_list <- list.dirs(macrosheds_root, recursive = F, full.names = F)
```

Then remove any domains outside of the CONUS.

```{r}
domain_list <- domain_list[!domain_list %in% c('arctic', 'mcmurdo', 'krycklan', 'luquillo')]
```


Make list of all site files.

```{r}
site_list <- as.character()
for(i in domain_list){
    loop_sites <- list.files(here('ms_data', i, 'stream_chemistry'),
                                       pattern = '.feather$', full.names = T)
    site_list <- c(site_list,loop_sites)
    }
```

## Intro case at HJ Andrews

Load data for a single site from HJ Andrews to show workflow

```{r}
site_path <- site_list[50]
solute <- 'DOC'

prep_con_data <- function(site_path, solute){
    target_sol <- paste0('GN_', solute)
    read_feather(site_path) %>%
        filter(var == target_sol) %>%
        select(site_code, datetime, val) %>%
        na.omit()}
con_prep <- prep_con_data(site_path, solute)
```

Summarize to annual mean data and filter out years made from less than 12 observations

```{r}
nrow(con_prep)
 
make_mean_annual <- function(con_prep){con_prep %>%
        mutate(wy = as.integer(as.character(water_year(datetime, origin = 'usgs')))) %>%
        group_by(wy) %>%
        summarize(n = n(),
                  val = mean(val)) %>%
    filter(n > 11)}

con_ann <- make_mean_annual(con_prep)

nrow(con_ann)
```
Plot data

```{r}
plot_annual <- function(con_ann, solute){ggplot(con_ann, aes(x = wy, y = val))+
    geom_point() +
        labs(y = paste0('Mean annual ', solute))+
    theme_few()}

plot_ann <- plot_annual(con_ann, solute)
plot_ann
```

Now detect trends and add to the plot

```{r}
ss <- sens.slope(con_ann$val)
ss

sig_check <- ss$p.value < 0.05

slope_est <- signif(ss$estimates[[1]], digits = 3) 

plot_ann <- plot_ann +
                labs(caption = paste("Sen's slope of", slope_est,
                     'from', nrow(con_ann), 'years of data.'))

plot_ann
```

## Make national DOC map

Now apply to all sites in the list

```{r}
ingest_national <- function(site_list, solute, data_type){
out_list <- list()
out_res <- tibble(site_code = as.character(),
                    slope = as.numeric(),
                    p = as.numeric(),
                    n = as.numeric())
out_data <- tibble(site_code = as.character(),
                   wy = as.numeric(),
                   n = as.numeric(),
                   doc = as.numeric())
for(i in 1:length(site_list)){
# load  and prep data
site_path <- site_list[i]

if(data_type == 'con'){con_prep <- prep_con_data(site_path, solute)}
else{ con_prep <- prep_flux_data(site_path, solute)}
site <-  con_prep$site_code[1]


# summarize data
if(data_type == 'con'){
    if(nrow(con_prep)>4){
    con_ann <- make_mean_annual(con_prep)
    }else{next}
}else{
   if(nrow(con_prep)>3){
    con_ann <- con_prep
    }else{next} 
}

if(nrow(con_ann) > 2){
# check for trends
ss <- sens.slope(con_ann$val)
ss

slope_est <- signif(ss$estimates[[1]], digits = 3)

loop_res <- tibble(site_code = site,
                    slope = ss$estimates[[1]],
                    p = ss$p.value[[1]],
                    n = nrow(con_ann))

loop_data <- con_ann %>%
    mutate(site_code = site)

}else{next}
#save to frames
out_data <- rbind(out_data, loop_data)
out_res <- rbind(out_res,loop_res)
}
out_list <- list(out_data, out_res)
return(out_list)
} # end function

solute <- 'DOC'
out_list <- ingest_national(site_list, solute, 'con')
out_res <- out_list[[2]]
out_data <- out_list[[1]]

```

Make functions to plot and map results

```{r}
# make plotter for popups
sc_plotter <- function(site){

  data <-filter(out_data, site_code == site)
  slope_est <- signif(out_res$slope[out_res$site_code == site], digits = 2)
  p <- signif(out_res$p[out_res$site_code == site], digits = 2)

  p <- plot_annual(data, solute) +
                labs(caption = paste("Sen's slope of", slope_est,
                     'from', nrow(data), 'years of data. p =', p),
                     title = site)
  return(p)
}

# make map data
make_map_data <- function(out_res){
inner <- out_res %>%
    mutate(significant = p < 0.05,
           status = 0) %>%
    left_join(site_data, by = 'site_code') %>%
    filter(!is.na(longitude), !is.na(latitude)) %>%
    st_as_sf(., coords = c('longitude', 'latitude'), crs = 4326)

inner$status[inner$significant == 1 & inner$slope > 0] <- 'Increasing'
inner$status[inner$significant == 0] <- 'No sig. trend'
inner$status[inner$significant == 1 & inner$slope < 0] <- 'Decreasing'
inner
}

map_data <- make_map_data(out_res)

# set color ramps
# colors = colorRampPalette(c('blue', 'gray', 'darkred'))

# apply mapping and plotting
make_map <- function(map_data){
p_all <- lapply(map_data$site_code, sc_plotter)

map_data <- map_data %>%
    mutate(mean_plots = map(site_code, sc_plotter))

mapview(map_data,
               zcol = 'status',
               cex = 'n',
               label = 'domain',
               col.regions = c('blue', 'darkred', 'gray'),
               popup = leafpop::popupGraph(map_data$mean_plots),
               layer.name = paste0(solute, ' Trend'))
}
```

Make the map!

```{r}
m1 <- make_map(map_data)
m1
```

## Apply to sulfate

```{r}
solute <- 'SO4_S'
out_list_so4 <- ingest_national(site_list, solute, 'con')
out_res <- out_list_so4[[2]]
out_data <- out_list_so4[[1]]

m2 <- out_res %>%
    make_map_data()%>%
    make_map()

m2
```

## Apply to pH

```{r}
solute <- 'pH'
out_list_ph <- ingest_national(site_list, solute, 'con')
out_res <- out_list_ph[[2]]
out_data <- out_list_ph[[1]]

m3 <- out_res %>%
    make_map_data()%>%
    make_map()

m3
```

# Loads

Now to repeat with loads. Making a function ('prep_flux_data') to make flux data work with previous functions. 

```{r}

domain_list <- list.dirs(here('ms_flux_annual'), recursive = F, full.names = F)

domain_list <- domain_list[!domain_list %in% c('arctic', 'mcmurdo', 'krycklan', 'luquillo')]

site_list <- as.character()
for(i in domain_list){
    loop_sites <- list.files(here('ms_flux_annual', i, 'stream_flux'),
                                       pattern = '.feather$', full.names = T)
    site_list <- c(site_list,loop_sites)
}

prep_flux_data <- function(site_path, solute){
    target_sol <- paste0('GN_', solute)
    
    data <- read_feather(site_path) %>%
        filter(var == target_sol) %>%
        na.omit()
    
    if(nrow(data > 1)){
    data <- data %>%
        dplyr::filter(ms_recommended == 1) %>%
        select(site_code, wy, val) %>%
        na.omit()}
    
    data
    
    }

for(i in 1:length(site_list)){
site_path <- site_list[i]
prep_flux_data(site_path, 'DOC')
}
```

Now generate maps. 

## DOC Loads

```{r}
solute <- 'DOC'
out_list <- ingest_national(site_list, solute, 'flux')
out_res <- out_list[[2]]
out_data <- out_list[[1]]

m4 <- out_res %>%
    make_map_data()%>%
    make_map()

m4
```

## Sulfate Loads

```{r}
solute <- 'SO4_S'
out_list <- ingest_national(site_list, solute, 'flux')
out_res <- out_list[[2]]
out_data <- out_list[[1]]

m5 <- out_res %>%
    make_map_data()%>%
    make_map()

m5
```

