---
title: "MacroSheds Solute Trends"
author: "Nick Gubbins"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)


library(here)
library(tidyverse)
library(lubridate)

source('analyze_trend.R')

```

Scope: provide a high level analysis of trends in riverine solute concentrations throughout the macrosheds database. 

After running load_data, you can run the following functions to generate the trends data set.

```{r, include=T}
#analyze_trend(target_solute = 'SO4_S', ms_root = here('ms_data'))
#analyze_trend(target_solute = 'pH', ms_root = here('ms_data'))
#analyze_trend(target_solute = 'DOC', ms_root = here('ms_data'))
```

The 'analyze_trend' function calculates annual and seasonal trends for sites that meet the required data cutoffs. The current cutoffs are

- Site is in the CONUS

- Data across at least 10 water years

- Data across at least 3 seasons

- There are at least 3 samples in 10 or more seasons




