---
title: "MacroSheds Solute Trends"
author: "Nick Gubbins"
output: html_document
toc: TRUE
theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      include = TRUE,
                      tidy.opts = list(width.cutoff = 60), 
                      #tidy = TRUE,
                      warning = F,
                      message = F)
# needed packages
library(here)
library(tidyverse)
library(lubridate)
library(mapview)
mapviewOptions(fgb = FALSE, georaster = FALSE)
library(sf)
library(tigris)
library(RColorBrewer)
library(leaflet)
library(sp)
library(ggthemes)
library(macrosheds)
library(tidyr)

# source model fitting function
source('analyze_trend.R')

# set color ramps
colors = colorRampPalette(c('blue', 'dark red'))

```

# Scope
Provide a high level analysis of trends in riverine solute concentrations throughout the MS database. 

# Calculate Trends
After running load_data, you can run the following functions to generate the trends data set. The function will apply to all the sites you have downloaded in your MS data directory.

```{r}
#analyze_trend(target_solute = 'SO4_S', ms_root = here('ms_data'))
#analyze_trend(target_solute = 'pH', ms_root = here('ms_data'))
#analyze_trend(target_solute = 'DOC', ms_root = here('ms_data'))
```

The 'analyze_trend' function calculates annual and seasonal trends for sites that meet the required data cutoffs. The current cutoffs are

- Site is in the CONUS

- Data across at least 10 water years

- Data across at least 3 seasons

- There are at least 3 samples in 10 or more seasons

The function then applies 5 linear models (annual + four seasons) across mean, low (5% percentile value), and high concentrations for each time period. The function uses USGS water years for annual aggregation and simple, three month periods for each season.

# Summarizing trends

```{r}
trends <- read_csv(here('trends','SO4_S_trends.csv'), show_col_types = FALSE) %>%
    rbind(.,read_csv(here('trends','DOC_trends.csv'), show_col_types = FALSE)) %>%
    rbind(.,read_csv(here('trends','pH_trends.csv'), show_col_types = FALSE)) 

site_data <- ms_load_sites()

sf_trends <- trends %>%
    left_join(., site_data, by = c('site' = 'site_code', 'domain')) %>%
    filter(!is.na(longitude),
           !is.na(latitude)) %>%
    st_as_sf(., coords = c('longitude', 'latitude'), crs = 4326)

sf_trends %>%
    select(domain_fullname, site, n_ann, geometry) %>%
    distinct() %>%
    mapview(.,
            zcol = 'n_ann',
            legend = T,
            label = 'domain_fullname',
            layer.name = 'Years of data')
```


The map above shows good national coverage. Sites are colored by years of data available. 


```{r,include=F}
visulaize_trends <- function(solute, metric, annual, significant_only){

    target_solute <- solute #whatever you have data on
    target_annual <- annual # 1 or 0
    sig_check <- significant_only # 1 or 0
    target_metric <- metric #high, low, mean

target_trends <- sf_trends %>%
        filter(solute == target_solute) 
    
if(target_annual == 1){
    # for annual cases
    # get data together 
    target_trends <- target_trends %>%
        select(domain_fullname, site,
               contains(paste0('_ann_',target_metric)),
               geometry)
    
    #remove suffix
    colnames(target_trends) <- gsub(paste0('_',target_metric),'',colnames(target_trends))
    
    # check and filter for sig only trends
    if(significant_only == 1){target_trends <- filter(target_trends, p_ann < 0.05)}
    else{target_trends <- target_trends}
    
    # make plot
    trend_plot <- ggplot(target_trends)+
        geom_histogram(aes(x = slope_ann)) +
        labs(x = 'Slope of trendline',
             title = paste(target_metric, '', target_solute)) +
        theme_few()
    
    # add caption to plot + show if sig or all
    if(significant_only == 1){trend_plot<- trend_plot + labs(caption = paste('n=',nrow(target_trends),'(signifcant only)'))}
    else{trend_plot <- trend_plot + labs(caption = paste('n=',nrow(target_trends)))}
    
    # make map
    ann_map <- mapview(target_trends,
            zcol = 'slope_ann',
            legend = T,
            label = 'domain_fullname',
            layer.name = 'Trend',
            col.regions= colors)
    
    plot_list <- list(trend_plot, ann_map)
    return(plot_list)
    } # end annual if statement
    else{
    
    # grab seasonal data
    target_trends <- target_trends %>%
        select(domain_fullname, site,
               contains(paste0('_f_',target_metric)),
               contains(paste0('_su_',target_metric)),
               contains(paste0('_sp_',target_metric)),
               contains(paste0('_w_',target_metric)),
               geometry)
    
    # drop suffix
    colnames(target_trends) <- gsub(paste0('_',target_metric),'',colnames(target_trends))
    
    
    # pivot into sensible format for graphing
    target_trends <- target_trends %>%
        # Stack metric/season info
        pivot_longer(names_to = "metric_season", values_to = "value",
                     cols = contains(c("p_", 'rsquared_', 'slope_'))) %>%
        # Separate metric and season
        separate(sep = "_", col = metric_season, into = c("metric", "season")) %>%
        # Create separate cols for each metric
        pivot_wider(names_from = "metric", values_from = "value")

    #check and filter for sig trends only
    if(significant_only == 1){target_trends <- filter(target_trends, p < 0.05)}
    else{target_trends <- target_trends}
    
    # make plot
    trend_plot <- ggplot(target_trends)+
        geom_histogram(aes(x = slope)) +
        labs(x = 'Slope of trendline',
             title = paste(target_metric, '', target_solute)) +
        facet_wrap(~season)+
        theme_few()
    
    # add caption to plot + show if sig or all
    if(significant_only == 1){trend_plot<- trend_plot + labs(caption = paste('n=',nrow(target_trends),'(signifcant only)'))}
    else{trend_plot <- trend_plot + labs(caption = paste('n=',nrow(target_trends)))}

    # make maps
    f_map <- target_trends %>%
        filter(season == 'f') %>%
        st_as_sf() %>%
    mapview(.,
            zcol = 'slope',
            legend = T,
            label = 'domain_fullname',
            layer.name = 'Fall Trend',
            col.regions= colors)
    

    w_map <- target_trends %>%
        filter(season == 'w') %>%
        st_as_sf() %>%
    mapview(.,
            zcol = 'slope',
            legend = T,
            label = 'domain_fullname',
            layer.name = 'Winter Trend',
            col.regions= colors)
    
    sp_map <- target_trends %>%
        filter(season == 'sp') %>%
        st_as_sf() %>%
    mapview(.,
            zcol = 'slope',
            legend = T,
            label = 'domain_fullname',
            layer.name = 'Spring Trend',
            col.regions= colors)
    
    su_map <- target_trends %>%
        filter(season == 'su') %>%
        st_as_sf() %>%
    mapview(.,
            zcol = 'slope',
            legend = T,
            label = 'domain_fullname',
            layer.name = 'Summer Trend',
            col.regions= colors)
    plot_list <- list(trend_plot, f_map, w_map, sp_map, su_map)
    return(plot_list)
            # print(paste('Fall', target_solute, '(', target_metric, ')')),
            # f_map,
            # print(paste('Winter', target_solute, '(', target_metric, ')')),
            # w_map,
            # print(paste('Spring', target_solute, '(', target_metric, ')')),
            # sp_map,
            # print(paste('Summer', target_solute, '(', target_metric, ')')),
            # su_map
    #)
    } #end seasonal else
}# end function

```

# Annual DOC Trends

## Mean

```{r, echo = F}
plots <- visulaize_trends(solute = 'DOC', metric = 'mean', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

## High

```{r, echo = F}
plots <- visulaize_trends(solute = 'DOC', metric = 'high', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

## Low

```{r, echo = F}
plots <- visulaize_trends(solute = 'DOC', metric = 'low', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

# Seasonal DOC Trends

## Mean

```{r, echo=F}
plots <- visulaize_trends(solute = 'DOC', metric = 'mean', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```

## High

```{r, echo=F}
plots <- visulaize_trends(solute = 'DOC', metric = 'high', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```


## Low

```{r, echo=F}
plots <- visulaize_trends(solute = 'DOC', metric = 'low', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```

# Annual pH Trends

## Mean

```{r, echo = F}
plots <- visulaize_trends(solute = 'pH', metric = 'mean', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

## High

```{r, echo = F}
plots <- visulaize_trends(solute = 'pH', metric = 'high', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

## Low

```{r, echo = F}
plots <- visulaize_trends(solute = 'pH', metric = 'low', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

# Seasonal pH Trends

## Mean

```{r, echo=F}
plots <- visulaize_trends(solute = 'pH', metric = 'mean', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```

## High

```{r, echo=F}
plots <- visulaize_trends(solute = 'pH', metric = 'high', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```


## Low

```{r, echo=F}
plots <- visulaize_trends(solute = 'pH', metric = 'low', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```


# Annual SO4_S Trends

## Mean

```{r, echo = F}
plots <- visulaize_trends(solute = 'SO4_S', metric = 'mean', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

## High

```{r, echo = F}
plots <- visulaize_trends(solute = 'SO4_S', metric = 'high', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

## Low

```{r, echo = F}
plots <- visulaize_trends(solute = 'SO4_S', metric = 'low', 
                 annual = 1, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
```

# Seasonal SO4_S Trends

## Mean

```{r, echo=F}
plots <- visulaize_trends(solute = 'SO4_S', metric = 'mean', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```

## High

```{r, echo=F}
plots <- visulaize_trends(solute = 'SO4_S', metric = 'high', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```


## Low

```{r, echo=F}
plots <- visulaize_trends(solute = 'SO4_S', metric = 'low', 
                 annual = 0, significant_only = 1)

plots[1][[1]]
plots[2][[1]]
plots[3][[1]]
plots[4][[1]]
plots[5][[1]]
```
